---
title: "Details of the PackageR Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{package_details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Description

`PackageR` package reads and plots Fatality Analysis Reporting System (FARS) data 
from the [National Highway Traffic Safety 
Administration](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars).

[FARS](https://en.wikipedia.org/wiki/Fatality_Analysis_Reporting_System) is a
nationwide census, providing the American public yearly data, regarding fatal
injuries suffered in motor vehicle traffic crashes.


## Load Packages

```{r, echo = FALSE, include = FALSE}
library(PackageR)
library(dplyr)
library(maps)
```

## Package functions

There are three exported functions available to users:

- `fars_read()`
- `fars_summarize_years()`
- `fars_map_state()`

Further we give a short description with examples how to use the functions. For the purposes of these examples we will use data from FARS, years 2013-2015, that can be found in the package directory under `\extdata` folder.

## Read FARS data files

The function `fars_read()` provides a way how to read data from a file path provided as an argument. It returns a `tibble` with loaded data:

```{r fars_read_example}
filename <- system.file("extdata/accident_2013.csv.bz2", package = "PackageR")
fars_read(filename)
```

## Summarize number of accidents

The next function `fars_summarize_years()` takes a vector or list of years (numeric/integer values) as an argument and if the corresponding files with data are available, it returns a pivot table with number of accidents per year and month. The data files need to be located in the working directory.

```{r fars_summarize_years_example}
setwd(system.file("extdata", package = "PackageR"))
fars_summarize_years(2013:2015)
```

## Plot accident locations

The last function available in the package, `fars_map_state()` takes a state index and year as arguments and returns a plot of the state with accident locations. The state indices need to correspond available indeces from the data set. Here are some examples:

```{r fars_map_state_example}
setwd(system.file("extdata", package = "PackageR"))
fars_map_state(45, 2015)
```

# Links:
* [Package PackageR GitHub](https://github.com/ricardomartins2/PackageR)
* [Package PackageR Travis CI](https://travis-ci.org/ricardomartins2/PackageR)

